Supplementary Information

### 2021-02-11 doc/manuscript/MEE_manuscript.Rmd
put all climwin plots in SI =/
#91

#SI figure numbers
all_recruitment_plots="S1-S10"

@@

all_climwin = "S15-S17"
climwin_sameV_sameW = "S15"
climwin_sameV_diffW = "S16"
climwin_diffV_diffW = "S17"

all_GLS = "S18-S37"
GLS_BCNM = "S18"
GLS_BCNM_year = "S19"
GLS_HKK = "S20"
GLS_HKK_year = "S21"
GLS_SCBI = "S22"
GLS_SCBI_year = "S23"
GLS_LDW = "S24"
GLS_LDW_year = "S25"
GLS_HF = "S36"
GLS_HF_year = "S27"
GLS_ZOF = "S28"
GLS_ZOF_year = "S29"
GLS_NIO = "S30"
GLS_NIO_year = "S31"
GLS_LT = "S32"
GLS_LT_year = "S33"
GLS_CB = "S34"
GLS_CB_year = "S35"
GLS_SC = "S36"
GLS_SC_year = "S37"

SC_1970break = "S38"

RW_interactions_all= "S39"
BAI_interactions_all= "S40"

decadal_analysis_figs_all = "S41-S50"
decadal_plot_BCNM = "S41"
decadal_plot_HKK = "S42"
decadal_plot_SCBI = "S43"
decadal_plot_LDW = "S44"
decadal_plot_HF = "S45"
decadal_plot_ZOF = "S46"
decadal_plot_NIO = "S47"
decadal_plot_LT = "S48"
decadal_plot_CB = "S49"
decadal_plot_SC = "S50"

#SI figure captions
recruitment_plot_BCNM_legend = "Figure S1. (recruitment BCNM)"
recruitment_plot_HKK_legend = "Figure S2. (recruitment HKK)"
recruitment_plot_SCBI_legend = "Figure S3. (recruitment SCBI)"
recruitment_plot_LDW_legend = "Figure S4. (recruitment LDW)"
recruitment_plot_HF_legend = "Figure S5. (recruitment HF)"
recruitment_plot_ZOF_legend = "Figure S6. (recruitment ZOF)"
recruitment_plot_NIO_legend = "Figure S7. (recruitment NIO)"
recruitment_plot_LT_legend = "Figure S8. (recruitment LT)"
recruitment_plot_CB_legend = "Figure S9. (recruitment CB)"
recruitment_plot_SC_legend = "Figure S10. (recruitment SC)"

@@

climwin_sameV_sameW_legend = "Figure S15. (PRE at SCBI)"
climwin_sameV_diffW_legend = "Figure S16. (PET at SCBI)"
climwin_diffV_diffW_legend = "Figure S17. (TMX/TMP at HKK)"

GLS_BCNM_legend = "Figure S18. Best GLS models including climate and DBH for Barro Colorado Nature Monument (Panama)"
GLS_BCNM_year_legend = "Figure S19. Best GLS models including climate, DBH, and year for Barro Colorado Nature Monument (Panama)"
GLS_HKK_legend = "Figure S20. Best GLS models including climate and DBH for Huai Kha Khaeng (Thailand)"
GLS_HKK_year_legend = "Figure S21. Best GLS models including climate, DBH, and year for Huai Kha Khaeng (Thailand)"
GLS_SCBI_legend = "Figure S22. Best GLS models including climate and DBH for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_SCBI_year_legend = "Figure S23. Best GLS models including climate, DBH, and year for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_LDW_legend = "Figure S24. Best GLS models including climate and DBH for Lilley Dickey Woods (Indiana, USA)"
GLS_LDW_year_legend = "Figure S25. Best GLS models including climate, DBH, and year for Lilley Dickey Woods (Indiana, USA)"
GLS_HF_legend = "Figure S26. Best GLS models including climate and DBH for Harvard Forest (Massachusetts, USA) "
GLS_HF_year_legend = "Figure S27. Best GLS models including climate, DBH, and year for Harvard Forest (Massachusetts, USA) "
GLS_ZOF_legend = "Figure S28. Best GLS models including climate and DBH for Zofin Forest (Czech Republic)"
GLS_ZOF_year_legend = "Figure S29. Best GLS models including climate, DBH, and year for Zofin Forest (Czech Republic)"
GLS_NIO_legend = "Figure S30. Best GLS models including climate and DBH for Niobrara (Nebraska, USA)"
GLS_NIO_year_legend = "Figure S31. Best GLS models including climate, DBH, and year for Niobrara (Nebraska, USA)"
GLS_LT_legend = "Figure S32. Best GLS models including climate and DBH for Little Tesuque (New Mexico, USA)"
GLS_LT_year_legend = "Figure S33. Best GLS models including climate, DBH, and year for Little Tesuque (New Mexico, USA)"
GLS_CB_legend = "Figure S34. Best GLS models including climate and DBH for Cedar Breaks (Utah, USA)"
GLS_CB_year_legend = "Figure S35. Best GLS models including climate, DBH, and year for Cedar Breaks (Utah, USA)"
GLS_SC_legend = "Figure S36. Best GLS models including climate and DBH for Scotty Creek (Northwest Territory, Canada)"
GLS_SC_year_legend = "Figure S37. Best GLS models including climate, DBH, and year for Scotty Creek (Northwest Territory, Canada)"

SC_1970break_legend = "Figure S38. Climate responses at Scotty Creek (Northwest Territory, Canada) before and after 1970."

RW_interactions_all_legend = "Figure S39. (RW_interactions_all)"
BAI_interactions_all_legend ="Figure S40. (BAI_interactions_all)"

decadal_plot_BCNM_legend = "Figure S41. (Decadal BCNM)"
decadal_plot_HKK_legend = "Figure S42. (Decadal HKK)"
decadal_plot_SCBI_legend = "Figure S43. (Decadal SCBI)"
decadal_plot_LDW_legend = "Figure S44. (Decadal LDW)"
decadal_plot_HF_legend = "Figure S45. (Decadal HF)"
decadal_plot_ZOF_legend = "Figure S46. (Decadal ZOF)"
decadal_plot_NIO_legend = "Figure S47. (Decadal NIO)"
decadal_plot_LT_legend = "Figure S48. (Decadal LT)"
decadal_plot_CB_legend = "Figure S49. (Decadal CB)"
decadal_plot_SC_legend = "Figure S50. (Decadal SC)"

@@

```{r , eval=FALSE, echo=FALSE}
## Tallies for the following paragraph (Feb 11)
#sameV_sameW: LDWp, NMt,NMp, NIOt, NIOp, SCBIp,  ZOFp,ZOFt,  (8)
#sameV_diffW: BCNMp, BCNMt, HFt, HKKp, LDWt,  SCBIt, SCp,SCt,  CBp, CBt (10)
#diffV_diffW: HFp, HKKt
```
The three metrics of growth ($RW$, $BAI$, and $\Delta AGB$)  exhibited similar strength of correlation and direction of response to climate variables across the range of potential time windows, but the optimal time window or even the top climate variable sometimes differed (Figs. `r all_climwin`).
In 8 of 20 cases, both the optimal climate variable and time window were identical across growth metrics (e.g., Fig. `r climwin_sameV_sameW`). 
In 10 cases, *climwin* identified the same climate variable but different (often overlapping) time windows (e.g., Fig. `r climwin_sameV_diffW`). 
Finally, in two cases of variables that had only weak effects and mixed responses among species in the final models (temperature variable group at HKK, precipitation variable group at HF; Figs. `r clim_sens`, `r GLS_HKK`, `r GLS_HF`), *climwin* identified different climate variables and different time windows (e.g., Fig. `r climwin_diffV_diffW`). 
Henceforth, unless otherwise noted, we focus on the climate sensitivities identified with $RW$ as the growth metric and for the full set of cores (*i.e.*, including those for which $DBH$ could not be reconstructed).

@@

A temperature variable was included in the best model for all but eight site-species combinations, with predominantly negative responses, particularly at the higher end of the temperature range (`r percent_temperature_responses_1st_order_negative + percent_temperature_responses_1st_order_pos_2nd_order_neg`%; `r percent_temperature_responses_1st_order_negative`% with negative first-order term, `r percent_temperature_responses_1st_order_pos_2nd_order_neg`% with positive first-order term but negative second-order term; Fig. `r clim_sens`).
The most commonly selected variables within the temperature group were $T_{max}$ and $PET$, which were identified by *climwin* as the top temperature-related driver at five and three of the ten sites, respectively, noting that $PET$ was not available for two sites. 
$T_{mean}$ and $T_{min}$ were each identified as the top driver at only one site (NIO and BCNM, respectively), noting that the effects of $T_{min}$ at BCNM were only marginally significant (Fig. `r clim_sens`).
Optimal time windows most commonly occurred during the peak growing season of the current year (n= 5 of 10 sites), but there were cases where optimal windows occurred during the preceding dry season (BCNM), late winter/early spring (HF, ZOF), or the previous growing season (NIO, CB).
Within the tropics, there was minimal effect of temperature at BCNM and a negative effect of wet season $T_{max}$ for three of four species at HKK.
For temperate sites with optimal time windows covering the current and/or past growing season, responses were universally negative (*i.e.*, negative first-order linear or unimodal, peaking at temperatures lower than the long-term mean). 
In contrast, there were positive effects of Jan-March $T_{max}$ for all three species at ZOF and of March $PET$ for *Tsuga Canadensis* at HF, the latter contrasting with a negative response of the other three species at HF (Fig. `r clim_sens`).
At the highest-latitude site (SC), which has undergone rapid warming and permafrost melt, *Picea mariana* responded positively (but with wide 95% CI on the slope) to temperature over the full analysis period (1903-2013); however, responses were predominately positive prior to 1970 and predominantly negative afterwards (Fig. `r SC_1970break`).

>>>>

#SI figure numbers
#SI figure numbers
all_recruitment_plots="S1-S10"

@@

all_climwin = "S15-S34"
climwin_BCNM_W = "S15"
climwin_BCNM_T = "S16"
climwin_HKK_W = "S17"
climwin_HKK_T = "S18"
climwin_SCBI_W = "S19"
climwin_SCBI_T = "S20"

all_GLS = "S35-S54"
GLS_BCNM = "S35"
GLS_BCNM_year = "S36"
GLS_HKK = "S37"
GLS_HKK_year = "S38"
GLS_SCBI = "S39"
GLS_SCBI_year = "S40"
GLS_LDW = "S41"
GLS_LDW_year = "S42"
GLS_HF = "S343"
GLS_HF_year = "S44"
GLS_ZOF = "S45"
GLS_ZOF_year = "S46"
GLS_NIO = "S47"
GLS_NIO_year = "S48"
GLS_LT = "S49"
GLS_LT_year = "S50"
GLS_CB = "S51"
GLS_CB_year = "S52"
GLS_SC = "S53"
GLS_SC_year = "S54"

SC_1970break = "S55"

RW_interactions_all= "S56"
BAI_interactions_all= "S57"

decadal_analysis_figs_all = "S58-S67"
decadal_plot_BCNM = "S58"
decadal_plot_HKK = "S59"
decadal_plot_SCBI = "S60"
decadal_plot_LDW = "S61"
decadal_plot_HF = "S62"
decadal_plot_ZOF = "S63"
decadal_plot_NIO = "S64"
decadal_plot_LT = "S65"
decadal_plot_CB = "S66"
decadal_plot_SC = "S67"

#SI figure captions
recruitment_plot_BCNM_legend = "Figure S1. Density plot of core record start years by species for Barro Colorado Nature Monument (Panama)."
recruitment_plot_HKK_legend = "Figure S2. Density plot of core record start years by species for Huai Kha Khaeng (Thailand)."
recruitment_plot_SCBI_legend = "Figure S3. Density plot of core record start years by species for the Smithsonian Conservation Biology Institute (Virginia, USA)."
recruitment_plot_LDW_legend = "Figure S4. Density plot of core record start years by species for Lilley Dickey Woods (Indiana, USA)."
recruitment_plot_HF_legend = "Figure S5. Density plot of core record start years by species for Harvard Forest (Massachusetts, USA)."
recruitment_plot_ZOF_legend = "Figure S6. Density plot of core record start years by species for Zofin Forest (Czech Republic)."
recruitment_plot_NIO_legend = "Figure S7. Density plot of core record start years by species for Niobrara (Nebraska, USA)."
recruitment_plot_LT_legend = "Figure S8. Density plot of core record start years by species for Little Tesuque (New Mexico, USA)."
recruitment_plot_CB_legend = "Figure S9. Density plot of core record start years by species for Cedar Breaks (Utah, USA)."
recruitment_plot_SC_legend = "Figure S10. Density plot of core record start years by species for Scotty Creek (Northwest Territory, Canada)."

@@

climwin_BCNM_W_legend = "Figure S15. Climwin output for water variable group at BCNM."
climwin_BCNM_T_legend = "Figure S16. Climwin output for temperature variable group at BCNM."
climwin_HKK_W_legend =  "Figure S17. Climwin output for water variable group at HKK."
climwin_HKK_T_legend = "Figure S18. Climwin output for temperature variable group at HKK."
climwin_SCBI_W_legend  = "Figure S19. Climwin output for water variable group at SCBI."
climwin_SCBI_T_legend = "Figure S20. Climwin output for temperature variable group at SCBI."
climwin_LDW_W_legend = "Figure S21. Climwin output for water variable group at LDW."
climwin_LDW_T_legend = "Figure S22. Climwin output for temperature variable group at LDW."
climwin_HF_W_legend =  "Figure S23. Climwin output for water variable group at HF."
climwin_HF_T_legend = "Figure S23. Climwin output for temperature variable group at HF."
climwin_ZOF_W_legend  = "Figure S25. Climwin output for water variable group at ZOF."
climwin_ZOF_T_legend = "Figure S26. Climwin output for temperature variable group at ZOF."
climwin_NIO_W_legend = "Figure S27. Climwin output for water variable group at NIO."
climwin_NIO_T_legend = "Figure S28. Climwin output for temperature variable group at NIO."
climwin_LT_W_legend =  "Figure S29. Climwin output for water variable group at LT."
climwin_LT_T_legend = "Figure S30. Climwin output for temperature variable group at LT."
climwin_CB_W_legend  = "Figure S31. Climwin output for water variable group at CB."
climwin_CB_T_legend = "Figure S32. Climwin output for temperature variable group at CB."
climwin_SC_W_legend  = "Figure S33. Climwin output for water variable group at SC."
climwin_SC_T_legend = "Figure S34. Climwin output for temperature variable group at SC."

GLS_BCNM_legend = "Figure S35. Best GLS models including climate and DBH for Barro Colorado Nature Monument (Panama)"
GLS_BCNM_year_legend = "Figure S36. Best GLS models including climate, DBH, and year for Barro Colorado Nature Monument (Panama)"
GLS_HKK_legend = "Figure S37. Best GLS models including climate and DBH for Huai Kha Khaeng (Thailand)"
GLS_HKK_year_legend = "Figure S38. Best GLS models including climate, DBH, and year for Huai Kha Khaeng (Thailand)"
GLS_SCBI_legend = "Figure S39. Best GLS models including climate and DBH for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_SCBI_year_legend = "Figure S40. Best GLS models including climate, DBH, and year for the Smithsonian Conservation Biology Institute (Virginia, USA)"
GLS_LDW_legend = "Figure S41. Best GLS models including climate and DBH for Lilley Dickey Woods (Indiana, USA)"
GLS_LDW_year_legend = "Figure S42. Best GLS models including climate, DBH, and year for Lilley Dickey Woods (Indiana, USA)"
GLS_HF_legend = "Figure S43. Best GLS models including climate and DBH for Harvard Forest (Massachusetts, USA)"
GLS_HF_year_legend = "Figure S44. Best GLS models including climate, DBH, and year for Harvard Forest (Massachusetts, USA) "
GLS_ZOF_legend = "Figure S45. Best GLS models including climate and DBH for Zofin Forest (Czech Republic)"
GLS_ZOF_year_legend = "Figure S46. Best GLS models including climate, DBH, and year for Zofin Forest (Czech Republic)"
GLS_NIO_legend = "Figure S47. Best GLS models including climate and DBH for Niobrara (Nebraska, USA)"
GLS_NIO_year_legend = "Figure S48. Best GLS models including climate, DBH, and year for Niobrara (Nebraska, USA)"
GLS_LT_legend = "Figure S49. Best GLS models including climate and DBH for Little Tesuque (New Mexico, USA)"
GLS_LT_year_legend = "Figure S50. Best GLS models including climate, DBH, and year for Little Tesuque (New Mexico, USA)"
GLS_CB_legend = "Figure S51. Best GLS models including climate and DBH for Cedar Breaks (Utah, USA)"
GLS_CB_year_legend = "Figure S52. Best GLS models including climate, DBH, and year for Cedar Breaks (Utah, USA)"
GLS_SC_legend = "Figure S53. Best GLS models including climate and DBH for Scotty Creek (Northwest Territory, Canada)"
GLS_SC_year_legend = "Figure S54. Best GLS models including climate, DBH, and year for Scotty Creek (Northwest Territory, Canada)"

SC_1970break_legend = "Figure S55. Climate responses at Scotty Creek (Northwest Territory, Canada) before and after 1970."

RW_interactions_all_legend = "Figure S56. (RW_interactions_all)"
BAI_interactions_all_legend ="Figure S57. (BAI_interactions_all)"

decadal_plot_BCNM_legend = "Figure S58. (Decadal BCNM)"
decadal_plot_HKK_legend = "Figure S59. (Decadal HKK)"
decadal_plot_SCBI_legend = "Figure S60. (Decadal SCBI)"
decadal_plot_LDW_legend = "Figure S61. (Decadal LDW)"
decadal_plot_HF_legend = "Figure S62. (Decadal HF)"
decadal_plot_ZOF_legend = "Figure S63. (Decadal ZOF)"
decadal_plot_NIO_legend = "Figure S64. (Decadal NIO)"
decadal_plot_LT_legend = "Figure S65. (Decadal LT)"
decadal_plot_CB_legend = "Figure S66. (Decadal CB)"
decadal_plot_SC_legend = "Figure S67. (Decadal SC)"

@@

```{r , eval=FALSE, echo=FALSE}
## Tallies for the following paragraph (Feb 11)
#sameV_sameW: LDWp, NMt,NMp, NIOt, NIOp, SCBIp,  ZOFp,ZOFt,  (8)
#sameV_diffW: BCNMp, BCNMt, HFt, HKKp, LDWt,  SCBIt, SCp,SCt,  CBp, CBt (10)
#diffV_diffW: HFp, HKKt
```
The three metrics of growth ($RW$, $BAI$, and $\Delta AGB$)  exhibited similar strength of correlation and direction of response to climate variables across the range of potential time windows, but the optimal time window or even the top climate variable sometimes differed (Figs. `r all_climwin`).
In 8 of 20 cases, both the optimal climate variable and time window were identical across growth metrics (e.g., Fig. `r climwin_SCBI_W`). 
In 10 cases, *climwin* identified the same climate variable but different (often overlapping) time windows (e.g., Fig. `r climwin_SCBI_T`). 
Finally, in two cases of variables that had only weak effects and mixed responses among species in the final models (temperature variable group at HKK, precipitation variable group at HF; Figs. `r clim_sens`, `r GLS_HKK`, `r GLS_HF`), *climwin* identified different climate variables and different time windows (e.g., Fig. `r climwin_HKK_T`). 
Henceforth, unless otherwise noted, we focus on the climate sensitivities identified with $RW$ as the growth metric and for the full set of cores (*i.e.*, including those for which $DBH$ could not be reconstructed).

@@

A temperature variable was included in the best model for all but eight site-species combinations, with predominantly negative responses, particularly at the higher end of the temperature range (`r round(percent_temperature_responses_1st_order_negative + percent_temperature_responses_1st_order_pos_2nd_order_neg)`%; `r round(percent_temperature_responses_1st_order_negative)`% with negative first-order term, `r round(percent_temperature_responses_1st_order_pos_2nd_order_neg)`% with positive first-order term but negative second-order term; Fig. `r clim_sens`).
The most commonly selected variables within the temperature group were $T_{max}$ and $PET$, which were identified by *climwin* as the top temperature-related driver at five and three of the ten sites, respectively, noting that $PET$ was not available for two sites. 
$T_{mean}$ and $T_{min}$ were each identified as the top driver at only one site (NIO and BCNM, respectively), noting that the effects of $T_{min}$ at BCNM were only marginally significant (Fig. `r clim_sens`).
Optimal time windows most commonly occurred during the peak growing season of the current year (n= 5 of 10 sites), but there were cases where optimal windows occurred during the preceding dry season (BCNM), late winter/early spring (HF, ZOF), or the previous growing season (NIO, CB).
Within the tropics, there was minimal effect of temperature at BCNM and a negative effect of wet season $T_{max}$ for three of four species at HKK.
For temperate sites with optimal time windows covering the current and/or past growing season, responses were universally negative (*i.e.*, negative first-order linear or unimodal, peaking at temperatures lower than the long-term mean). 
In contrast, there were positive effects of Jan-March $T_{max}$ for all three species at ZOF and of March $PET$ for *Tsuga Canadensis* at HF, the latter contrasting with a negative response of the other three species at HF (Fig. `r clim_sens`).
At the highest-latitude site (SC), which has undergone rapid warming and permafrost melt, *Picea mariana* responded positively (but with wide 95% CI on the slope) to temperature over the full analysis period (1903-2013); however, responses were predominately positive prior to 1970 and predominantly negative afterwards (Fig. `r SC_1970break`).
